---
title: "Mapping the Coronavirus Outbreak"
author: "Garrett Allen"
date: "01-28-2020"
output: 
  html_document:
    toc: true
    number_sections: false
    toc_float: true
    df_print: paged
---

```{r set-up, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,comment = "#>", highlight = TRUE,fig.align = "center")
```

# Getting started

1. Clone your repo `appex04-[github_name]` to create a new project
   in RStudio Cloud under the STA 199 class space.

2. Configure git
```{r config-git-ex, eval=FALSE}
library(usethis)
use_git_config(user.name="Garrett.Allen", user.email="gpa5@duke.edu")
```

# Packages and Data

We'll make use of the following packages.

```{r load-packages}
library(sf)
library(tidyverse)
```

There are two data files we will use: `world.shp`, `coronavirus.shp`. We'll
read each of these in with function `st_read()` and save them as `world` and
`virus`.

```{r}
world <- st_read("data/world.shp", quiet = TRUE)
virus <- st_read("data/coronavirus.shp", quiet = TRUE)
nc <- st_read(system.file("shape/nc.shp", package = "sf"), quiet = TRUE)
```

# Tasks

## Task 1

Take a look at objects `world` and `virus`. How many fields exist for each
object? What type of geometry is associated with each `sf` object?

```{r task-1-world}
world
```

In "world", there are 241 observations of 9 fields. For this chart, the geometry is multipolygon.
```{r task-1-virus}
virus
```

in "virus" there are 44 observations and also 9 fields. For this plot, the geometry is point. 

## Task 2

#### Part 1

Use object `world` to create a world map of the countries. You'll want to use
functions `ggplot()` and `geom_sf()`.

```{r task-2-p1}
ggplot(world) + 
  geom_sf()
```

#### Part 2

Build on your map from Part 1 so that the countries have a fill color 
associated with the population estimate. Variable `pop_est` is in millions.
Be sure to label your map.

```{r task-2-p2, fig.width = 12, fig.height = 12}
ggplot(world) + 
  geom_sf(aes(fill = pop_est),) + 
  labs(title = "World Population In Millions", fill = "Millions") + 
  scale_fill_gradient(low = "lightblue", high = "red") + 
  theme_bw(base_size = 30)
```

## Task 3

Filter `world` for the country "China". Save the result as `china`.

```{r task-3-china}
china <- world %>% 
  filter(name == "China")
```

Next, we'll filter object `virus` for confirmed cases in China and save the
result as `china_cv`. The code is given below.

```{r task-3-china-cv}
china_cv <- virus %>% 
  filter(cntry_r == "Mainland China", !is.na(confrmd))
```

- `cntry_r == "Mainland China"` filters rows so we only keep information on
  China
  
- `!is.na(confrmd)` filters rows for where there is not a missing confirmed
  case of coronavirus

## Task 4

Use the template provided to overlay the point locations of the coronavirus
in China with a map of China. Have the size of the points reflect the
number of confirmed cases. Refer to slides 31-32 as a guide.

```{r task-4, fig.width=9, fig.height=6}
ggplot(data = china) +
  geom_sf(fill = "#DE2910") + 
  geom_sf(data = china_cv, aes(size = confrmd), color = "yellow", show.legend = "point") + 
  labs(title = "Coronavirus Outbreak in China", subtitle = "As of Jan 25, 2020", size = "Confirmed Cases") + 
  theme_bw()
```

*Hints:*

- you'll need another `geom_sf()` layer
- set point colors as `color = "#FFDE00"`
- inside this second `geom_sf()` include `show.legend = "point"` to have the 
  legend show points rather than squares

# Submission

## Stage, commit and push

1. Stage your modified files.
2. Commit your changes with an informative message.
3. Push your changes to your GitHub repo.
4. Verify your files were updated on GitHub.

# Slide notes

# References

1. Simple Features for R vignette, https://r-spatial.github.io/sf/
